Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-01-08 komlos.tex

\title[Komols Complexity, Coresets, and Sketches in Machine Learning]{Komols Complexity, Coresets, and Sketches in Machine Learning}
\usepackage{times}
% Use \Name{Author Name} to specify the name.
% If the surname contains spaces, enclose the surname
% in braces, e.g. \Name{John {Smith Jones}} similarly
% if the name has a "von" part, e.g \Name{Jane {de Winter}}.
% If the first letter in the forenames is a diacritic
% enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

@@

\begin{definition}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$ uniformly. 
The Rademacher complexity of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
R_m(\F) =  \frac{1}{m}\E_{\sigma} \sup_{f \in \F}  \left| \sum_{i=1}^{m}\sigma_i f(x_i) \right|
\]
\end{definition}
%
In other words, the Radamacher complexity of a family function $\F$ measures how much a function in the family can skew the values when weighted with random signs. The Komlos complexity measures the ability of the family function to skew an adversarial set of signs.

@@

\noindent It well known that $R_m$ is a good measure for generalization when data is chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them. We claim that $K_m$ is the measure more suitable for \zk{I think we should remove `active learning' from the list} creating coresets, and for creating mergeable sketches.

@@

 \zk{I actually am still having problems with justifying the name Komlos complexity. It might as well be called Banaszczyk's complexity or simply discrepancy complexity. In particular if you could prove Komlos' conjecture it wouldn't affect any of our results, nor simplify our proofs.. Either way, below is my suggested motivation for the definition}
To understand the motivation of the the Komlos Complexity, consider the following informal explanation of Radamacher Complexity applied to ML problems. In this ML problem there exists a distribution over $(x,y)$ pairs where we aim to predict $y$ given $x$. Our goal is to find a predictor from a family of prediction functions that suffers the least loss. A predictor maps to a loss value on each $x,y$ pair so we can discuss the problem of finding a function $f$ from a family $\F$ minimizing $\E_{x,y} f(x,y)$ where $f$ is the loss associated with the corresponding predictor on the $x,y$ pairs. Having a low Radamacher complexity means that if we sample $m$ pairs, split them into two uniformly at random and find the function $f$ minimizing the loss over the first half, the loss over the second half is bounded. This translates to a generalization bound stating that the function $f$ minimizing the loss over a uniform random sample of $(x,y)$ pairs, will perform well on the entire distribution. In other words, the Radamacher complexity gives a guarantee for the distribution loss of the minimizer of a finite set of pairs chosen uniformly at random.

Coming back to Komlos Complexity, having the ability to choose the signs in a non-uniform way lets us choose a subset of pairs $(x,y)$ such that the minimizer $f$ over the subset is guaranteed to have (roughly) the same performance on the entire set. This allows us to obtain a guarantee for the distribution loss of a minimizer of a finite set of pairs chosen in an adaptive way to the family function $\F$. This set is in fact a core-set and the Komlos complexity helps us determine that possible size of a coreset. The tradeoff of considering a coreset vs.\ a random sample is that on one hand, it is clearly much simpler to obtain a random sample. On the other hand we will show in the following sections several examples for which a coreset can be significantly smaller than the random sample while maintaining the same guarantees. This will be done by showing that for a wide range of interesting problems in machine learning $K_m(\F) = o(R_m(\F))$.

%\begin{theorem}
%Let $F(a) = \sum_i a_i$ for $a \in A$ and let $K_m(A) = O(c/m^\gamma)$ for any constant $\gamma \in (0,1]$.
%There exist a coreset for $A$ such that $\tilde{F}(a) = \sum_{i} w_i a_i$, $\operatorname{supp}(w) \le m$ and 
%$$
%\forall \; a \in A \;\;\;  |F(a) - \tilde F(a)| \le \eps n
%$$
%\end{theorem}
%\begin{proof}
%By halving the points repeatedly \el{TODO}.
%\end{proof}
\subsection{General Framework for Coresets}
In what follows it will be convenient to represent the function family as a family of vectors representing the functions. The vector family will be denoted as $Q$, where we use $q$ to denote a single vector corresponding to a function. We a universal function $f$ such that $f(x,q)$ corresponds to the value of the function matching $q$ when applied to $x$. We will assume from hereon that for any $x,q$ $|f(x,q)| \leq 1$; this allows to simplify the presentation. Our results can extend to $|f(x,q)| \leq R$ and will have dependencies on the bound $R$. Achieving these bounds is a purely technical task that we defer for the sake of readability.

Consider the function $F(q) = \sum_{i=1}^{m} f(x_i,q)$ and the signed sum of error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Now consider $F_1(q) = F(q) + E(q) = \sum_{i=1}^{m} f(x_i,q)  + \sum_{i=1}^{m} \sigma_i f(x_i,q)  = 2 \sum_{i | \sigma_i=1} f(x_i,q)$ and similarly $F_{-1}(q)$. We have that both $F_1(q), F_{-1}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|F_1(q) - F(q)| =  |E(q)|$.
%
The above is true for any choice of signs $\sigma_i$, specifically, for those minimizing $\max_q | E(q)|$.
Note that we can select signs such that $|E(q)| \le m K_m(\F)$.
A function is sketch-able if $K_m(\F) = o(m)$.
Typically we see that $K_m(\F) = O(c/m)$ or $K_m(\F)= c/\sqrt{m}$ for some $c$ which does not depend on $m$.

\begin{fact}
For any function family $\F$ with a corresponding Komlos Complexity $K_m(\F) = O(c/m)$ there exists a coreset of size 
$O(c/\eps)$ whose error is at most $\eps n$.
\end{fact}
\begin{fact}
For any function family $\F$ with a corresponding Komlos Complexity $K_m(\F) = O(c/\sqrt{m})$ there exists a coreset of size 
$O(c^2/\eps^2)$ whose error is at most $\eps n$.
\end{fact}

\noindent Proving both facts is trivial. 
At step one, create a coreset of size at most $n/2$ of items of weight $2$ and incur an error of at most $c$ or $c \sqrt{n}$.
Then, we repeat the process and create a coreset of size $n/4$ of items of weight $4$. Here, you incur error of $2c$ or $2c\sqrt{n/2}  = c\sqrt{2n}$.
Note that the sum of errors is a geometric sequence that is asymptotically dominated by its last element. 
Halting the compression at $\Theta(c/\eps)$ or $\Theta(c^2/\eps^2)$ items respectively achieves the goal.

%\begin{theorem}
%Let $A$ be of the special form $a_i = f(x_i,q) \in [0,1]$ and $K_m(A) = O(c/\sqrt{m})$.
%There exist and streaming mergeable coreset for $A$ with cardinality at most $\ell = O(c\log^{3}\log(1/\delta)/\eps^2)$.
%That is, let $F(q) = \sum_i f(x_i,q)$. One can process a stream of item $x_1,\ldots,x_n$ and produce $\tilde{F}(q) = \sum_{i} w_i a_i$ with $\operatorname{supp}(w) \le \ell$ such that 
%$$
%\forall \; q \in Q \text{\;\;with probability\;\;} 1-\delta  \;\;\;  |F(q) - \tilde F(q)| \le \eps n
%$$
%\end{theorem}
%
%
%\begin{theorem}
%Let $A$ be of the special form $a_i = f(x_i,q) \in [0,1]$ and $K_m(A) = O(c/m)$.
%There exist and streaming mergeable coreset for $A$ with cardinality at most $\ell = O(c\log^{2}\log(1/\delta)/\eps)$.
%That is, let $F(q) = \sum_i f(x_i,q)$. One can process a stream of item $x_1,\ldots,x_n$ and produce $\tilde{F}(q) = \sum_{i} w_i a_i$ with $\operatorname{supp}(w) \le \ell$ such that 
%$$
%\forall \; q \in Q \text{\;\;with probability\;\;} 1-\delta  \;\;\;  |F(q) - \tilde F(q)| \le \eps n
%$$
%\end{theorem}

\subsection{General Framework for Sketching} \label{sec:sketch}

The high level idea is the following. 
Consider a stream of the data points $x_i$. 
We maintain a buffer that can hold up to $m$ points.
Once collecting $m$ data points it performs a \emph{compact} operation that outputs at most $m/2$ data points of weight $2$ representing the original $m$ inputs. The compact operation partitions the $m$ points into two sets whose corresponding sums are very close. 
This is made formal with the following definitions:

 %
In what follows, we prove the following.
\begin{theorem} \label{thm:streaming}
For any function $f$ with a corresponding Komlos Complexity $K_m = O(c/m)$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
$$O\left(c\log^2\left(\min(\eps n/c, \log(n/\delta)) \right)/\eps\right)$$ 
whose error is at most $\eps n$ and failure probability at most $\delta$. For $\delta=0$ this results in a deterministic algorithm. The algorithm must be given $\eps, \delta$ as input and does not require a priori knowledge of the stream length $n$.
\end{theorem}

\begin{theorem} \label{thm:streaming2}
For any function $f$ with a corresponding Komlos Complexity of $K_m = O(c/\sqrt{m})$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
$$O\left(c^2\log^3\left(\min(\eps^2 n/c, \log(n/\delta)) \right)/\eps^2\right)$$ 
whose error is at most $\eps n$ and failure probability at most $\delta$. For $\delta=0$ this results in a deterministic algorithm. The algorithm must be given $\eps, \delta$ as input and does not require a priori knowledge of the stream length $n$.
\end{theorem}
%
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(\sqrt{k})$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
%$O(\log^2(\log(1/\delta))/\eps^2)$ whose error is at most $\eps n$ and failure probability at most $\delta$.
%\end{fact}
%
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(1)$ there exists a fully-mergeable deterministic streaming coreset algorithms of size 
%$O(\log^2(n)/\eps)$ whose error is at most $\eps n$.
%\end{fact}
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(\sqrt{k})$ there exists a fully-mergeable deterministic streaming coreset algorithms of size 
%$O(\log^2(n)/\eps^2)$ whose error is at most $\eps n$.
%\end{fact}

@@

\section{A Simple Algorithm for Kernel Density Estimation}

From section \ref{sec:analytic} we know that the Komlos complexity of the Gaussian kernel is $K_m = O(\sqrt{d}/m)$. The result however is non-constructive in the sense that we are not aware of an algorithm that can obtain the $\sigma$ signs efficiently. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. We show that for any positive kernel $K_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section when $d > \sqrt{m}$. More importantly though, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = x_1,\ldots, x_n$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as 
$$ f(q) = \sum_{i=1}^{n} K(x_i,q) $$
Here, $K$ is a \emph{positive definite kernel} function, typically based on the distance between $x,y$. The most frequent examples include
$$ K(x,q) = \exp(- \|x-y\|_2^2/\lambda^2)\;\;\; K(x,q) = \exp(- \|x-q\|/\lambda) \; \mbox{and}\;\;\; K(x,y) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

@@

\begin{proof}
%Using our framework, we need to bound $K_m = \frac{1}{m}\sum_{i=1}^m \sigma_i K(x_i,q)$. 
For any kernel $K$ there exist a mapping $\phi: \R^d \to {\cal V}$ to an inner product space $\cal V$ such that 
$$ K(x,q) = \ip{\phi(x), \phi(q)} $$
Using this function $\phi$ our objective function becomes
\[
\sum_{i=1}^m \sigma_i K(x_i,q) = \sum_{i=1}^m \sigma_i \ip{\phi(x_i), \phi(q)} =  \ip{ \sum_{i=1}^m \sigma_i \phi(x_i), \phi(q)} \leq  \|\phi(q)\| \cdot \left\|  \sum_{i=1}^m \sigma_i \phi(x_i) \right\| %\leq \left\|  \sum_{i=1}^m \sigma_i \phi(x_i) \right\| 
\]
Since $\|\phi(q)\| \leq 1$ we reduced the problem to bounding the norm of $ \sum_{i=1}^m \sigma_i \phi(x_i) $.
%
%This upper bound allows us to find values for $\sigma_i$ which bound $K_n$ using a simple algorithm.
%Set $\sigma_1 = 1$. Then for $i=2,\ldots,m$ set $\sigma_i = -\operatorname{sign} (\sum_{j=1}^{i-1}\sigma_j  K(x_j, x_i))$.
We will show by induction on $i$ that 
$$\left\| \sum_{j=1}^i \sigma_j \phi(x_j) \right\|^2 \le \sum_{j=1}^i \left\|\phi(x_j)\right\|^2 \leq i$$
This is trivially true for $i=1$ since $\|\phi(x)\| \leq 1$. 
%Assume by induction that $\| \sum_{j=1}^{i-1} \sigma_j \phi(x_j)\|^2 \le \sum_{j=1}^{i-1} \|\phi(x_j)\|^2$.
Using our induction assumption we get
\begin{eqnarray*}
\left\| \sum_{j=1}^{i}\sigma_j \phi(x_j)\right\|^2 =& \left\|\sum_{j=1}^{i-1}\sigma_j \phi(x_j)\right\|^2 + \|\phi(x_i)\|^2 + 2\ip{ \sum_{j=1}^{i-1}\sigma_j \phi(x_j), \sigma_i \phi(x_i)} \leq & \text{induction assumption}\\
& \sum_{j=1}^{i-1} \|\phi(x_j)\|^2 + \|\phi(x_i)\|^2 + 2\sigma_i \sum_{j=1}^{i-1}\sigma_j K(x_j, x_i) = & \text{choice of } \sigma_i \\
& \sum_{j=1}^{i} \|\phi(x_j)\|^2 - 2|\sum_{j=1}^{i-1}\sigma_j K(x_j, x_i)| \le \sum_{j=1}^{i} \|\phi(x_j)\|^2 \\
\end{eqnarray*}
%\begin{eqnarray*}
%\left\| \sum_{j=1}^{i}\sigma_j \phi(x_j)\right\|^2 &=& \left\|\sum_{j=1}^{i-1}\sigma_j \phi(x_j)\right\|^2 + \|\phi(x_i)\|^2 + 2\ip{ \sum_{j=1}^{i-1}\sigma_j \phi(x_j), \sigma_i \phi(x_i)} \\
%&\le& \sum_{j=1}^{i-1} \|\phi(x_j)\|^2 + \|\phi(x_i)\|^2 + 2\sigma_i \sum_{j=1}^{i-1}\sigma_j K(x_j, x_i) \mbox{\;\;\; by the induction assumption}\\ 
%&=& \sum_{j=1}^{i} \|\phi(x_j)\|^2 - 2|\sum_{j=1}^{i-1}\sigma_j K(x_j, x_i)| \le \sum_{j=1}^{i} \|\phi(x_j)\|^2
%\end{eqnarray*}
%This completes the proof that $\| \sum_{j=1}^{m}\sigma_j \phi(x_j)\| \le \sqrt{\sum_{j=1}^{i} \|\phi(x_j)\|^2} \le \sqrt{m}$ for our choice of values for $\sigma$ and shows that the Komlos complexity of any positive kernel is $O(1/\sqrt{m})$. 
\end{proof}

@@

%\section{Linear Regression}

%\section{Optimal Design of Experiments}
%We discuss the problem of linear regression with mean squared error. In the classic setting we are given $n$ points $x_1,\ldots,x_n \in \R^d$ along with $n$ labels $y_1,\ldots,y_n \in \R$ and wish to find $w \in \R^d$ minimizing
%$$ \sum_i (w^Tx_i - y_i)^2 $$
%In the field of optimal design of experiments, the labels are not given with the data. Instead, the learner must choose in advance a subset of datapoints for which they query the labels. The subset is chosen to provide the least error in a generalization bound, assuming $(x_1,y_1),\ldots,(x_n,y_n)$ are i.i.d from an unknown distribution.
%
%The way this is done in optimal design of experiments is as follows. Let $X$ be the set of $n$ points and consider a subset $B \subseteq X$. The empirical risk minimizer (ERM) for $B$ is 
%$$ w_B = \left( \sum_{x \in B} xx^T \right)^{\dagger} y_B $$
%%It is known that $w_B$ is an unbiased estimator of $w^*$, the optimal regressor for the distribution, where the noise comes from the realizations of the $y_i$'s. It follows that to minimize 
%%$$ E_{x \sim {\cal D}} \ip{w-w^*,x}^2 $$
%%which is the variance of $w$ measured by the distribution of data points, it suffice to minimize the second moment
%%$$ E_{x \sim {\cal D}} \ip{w,x}^2  \approx w^T \left( \frac{1}{n} \sum_{i=1}^n x_ix_i^T \right) w $$
%
%
%One can show that for $w_B$,
%$$ \sum_i (w_B^Tx_i - y_i)^2 = C \cdot \text{Tr} \left(\left( \sum_{x \in B} xx^T \right)^{\dagger} \right) $$
%\zk{This only works when $\sum_{x \in B} xx^T$ is full rank, otherwise we need to introduce some regularization and then things get messy.}
%
%We can use the results above for optimal design in the case where $w$ is restricted to have a unit Euclidean norm. Something that is commonly practiced. Since we will solve the problem for the subset we choose in an optimal way for gaussian noise, we end up with an unbiased estimator for $w^*$. The error is the variance of that unbiased estimator and minimizing the variance for an unbiased estimator is the same as minimizing its second moment we can aim to minimize
%$$ F(w) = w^T \left( \frac{1}{n} \sum_{i=1}^n x_ix_i^T \right) w $$
%This leads to the exact same objective as in column subset selection. Meaning that an $\eps$ core-set for the $x$'s gives a guaranteed error for a subset in the optimal design setting. Specifically we get that we can obtain a core-set of size
%$$ \min(1/\eps^2, \sqrt{d}/\eps)$$
%For the case of $\eps < 1/\sqrt{d}$, which is an interesting case in linear regression, a random sample of $\sqrt{d}/\eps$ points will not give a generalization bound of $\eps$ (there are known lower bounds, http://jmlr.org/papers/volume16/shamir15a/shamir15a.pdf contradicting this), meaning that we have a provable non-trivial guarantee for active learning. I'm not sure what statements are known already. 
%
%In the optimal design setting, they require keeping the covariance matrix. We could come up with a hueristic or even sketching algorithm that avoids maintaining $d^2$ points.

%\section{Logistic Regression}
%The following paper claims no streaming algorithms or coresets are available \cite{DBLP:journals/corr/abs-1805-08571}.
%This paper claims to have done that \cite{DBLP:conf/nips/HugginsCB16}.
%
%In \cite{DBLP:journals/corr/abs-1802-01751} (see paragraph before Lemma 2.2) they manage to obtain a coreset (with equal weights) of size $\sqrt{d \log(1/\eps)}/\eps$ for any kernel with certain properties. They do so by considering a matrix whose rows correspond to all the vectors in an $\eps$-net over the space, and columns correspond to the $n$ data points. Calling that matrix $A$, $A_{q,x} = K(q,x)$. They bound the discrepancy of the matrix defined as $\min_s \|As\|_\infty$ where $s$ is a sign vector by using a bound 
%$$ \min_s \|As\|_\infty \leq \sqrt{\log(N)} \|A\|_{\gamma_2} $$
%Here, $N$ is the size of the $\eps$ net and the $\gamma_2$ norm is 
%$$ \|A\|_{\gamma_2} = \min_{B,C: A=BC^T} \text{max norm of a row of B or C}$$
%They use the fact that for psd kernel $K$, $\|A\|_{\gamma_2} \leq 1$ (in fact it is equal to 1). Now, back to our case. For any fixed power $k$ we know that $K(x,q)=\ip{x,q}^k$ is a psd kernel, hence its corresponding matrix $A_k$ has $\gamma_2$ norm bounded by $1$. Now, the logistic function can be written in a taylor expansion whose coefficients, in absolute value, sum to a constant $C$. It follows that the $A$ matrix for the logistic function has a $\gamma_2$ norm bounded by $C$. The same arguments as in  \cite{DBLP:journals/corr/abs-1802-01751} can now prove that its discrepancy is bounded by $\sqrt{\log(N)} \approx \sqrt{d\log(1/\eps)}$
%
%A useful fact about the $\gamma_2$ norm is that it is indeed a norm. This is proven in \cite{tomczak1989banach}. In \cite{matouvsek2014factorization} this is listed in Section 2, Proposition 2.2 along with other known properties of the $\gamma_2$ norm. 
%
%
%
%
%
%\section{Coreset for linear functions (based on $\gamma_2$ norm)}
%
%Let $x_1,\ldots,x_n \in R^d$ be and $y_1,\ldots,y_n$ be labels associated to the data points. Let $K$ be a psd kernel. Consider a problem of learning a linear function aiming to minimize the loss of 
%$$\min_q \sum_{i=1}^n L(K(x_i, q), y_i)$$
%
%Our objective is to find a core-set of the $n$ data points. That is, we aim to find a set $S \subseteq [n]$ such that for every $q$ it holds that 
%$$ \left| \sum_{i=1}^n L(K(x_i, q), y_i) - \frac{n}{|S|} \sum_{i \in S} L(K(x_i, q), y_i) \right| \leq \eps n $$
%
%For a function $f(q,x)$, let $A_f$ be a matrix whose rows are an $\eps$-net over the Euclidean unit ball of $R^d$, and columns correspond to the different $x_i$'s. The value of $A_f$ at an entry corresponding to $(q,x)$ if $f(q,x)$.
%
%\begin{definition}
%For a matrix $A$,
%$$ \|A\|_{\gamma_2} = \min_{B,C: A=BC^T} \|B\|_{2 \to \infty} \|C\|_{2 \to \infty}$$
%is the $\gamma_2$ norm of $A$. 
%\end{definition}
% \cite{tomczak1989banach} prove that this is indeed a norm, in particular that 
% $\|A_1+A_2\|_{\gamma_2} \leq \|A_1\|_{\gamma_2} + \|A_2\|_{\gamma_2}$. 
%
%
%\begin{lemma} \label{lem:powers}
%For $k \geq 0$ let $f_k(q,x) = \ip{q,x}^k$. If holds that $\|A_{f_k}\|_{\gamma_2} \leq 1$
%\end{lemma}
%\begin{proof}
%For $k=0$ the statement is obvious. For $k>0$, notice that $f_k$ is a kernel, and in particular there exists a mapping $\phi_k: \R^d \to \R^{d^k}$ such that
%$$ \ip{q,x}^k = \ip{\phi_k(q), \phi_k(x)} $$
%Furthermore, 
%$$ \|\phi_k(x)\| = \sqrt{\ip{\phi_k(x), \phi_k(x)}} = \sqrt{\|x\|^{2k}} = \|x\|^k \leq 1 $$
%Since all $x,q$ have at most unit norm we get a bound over the $\gamma_2$ norm by viewing the $B,C$ matrices obtained from the $\phi_k$ embeddings of the different $x,q$'s
%\end{proof}
%
%\begin{lemma}
%Let $L:\R \to \R$ be a function that can be written as 
%$$ L(z) = \sum_{k=0}^\infty \alpha_k z^k$$

>>>>

\title[Discrepancy, Coresets, and Sketches in Machine Learning]{Discrepancy, Coresets, and Sketches in Machine Learning}
\usepackage{times}
% Use \Name{Author Name} to specify the name.
% If the surname contains spaces, enclose the surname
% in braces, e.g. \Name{John {Smith Jones}} similarly
% if the name has a "von" part, e.g \Name{Jane {de Winter}}.
% If the first letter in the forenames is a diacritic
% enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

@@

\begin{definition}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$ uniformly at random. 
The Rademacher complexity of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
R_m(\F) =  \frac{1}{m}\E_{\sigma} \sup_{f \in \F}  \left| \sum_{i=1}^{m}\sigma_i f(x_i) \right|
\]
\end{definition}
%
In other words, the Radamacher complexity of a family function $\F$ measures how much a function in the family can skew the values when weighted with random signs. The Komlos complexity measures the ability of the family function to skew an adversarial set of signs.

@@

\noindent It well known that $R_m$ is a good measure for generalization when data is chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them. We claim that $K_m$ is the measure more suitable for creating coresets, and for creating mergeable sketches.

@@

 \zk{I actually am still having problems with justifying the name Komlos complexity. It might as well be called Banaszczyk's complexity or simply discrepancy complexity. In particular if you could prove Komlos' conjecture it wouldn't affect any of our results, nor simplify our proofs.. Either way, below is my suggested motivation for the definition}
 \el{You are right about the name. In fact, I've been thinking about it myself. The discrepancy we get is actually much lower than what Komlos conjecture says... So, I am happy with with calling it simple \it{the discrepancy.}}
To understand the motivation of the the Komlos Complexity, consider the following informal explanation of Radamacher Complexity applied to ML problems. In this ML problem there exists a distribution over $(x,y)$ pairs where we aim to predict $y$ given $x$. Our goal is to find a predictor from a family of prediction functions that suffers the least loss. A predictor maps to a loss value on each $x,y$ pair so we can discuss the problem of finding a function $f$ from a family $\F$ minimizing $\E_{x,y} f(x,y)$ where $f$ is the loss associated with the corresponding predictor on the $x,y$ pairs. Having a low Radamacher complexity means that if we sample $m$ pairs, split them into two uniformly at random and find the function $f$ minimizing the loss over the first half, the loss over the second half is bounded. This translates to a generalization bound stating that the function $f$ minimizing the loss over a uniform random sample of $(x,y)$ pairs, will perform well on the entire distribution. In other words, the Radamacher complexity gives a guarantee for the distribution loss of the minimizer of a finite set of pairs chosen uniformly at random.

Coming back to Komlos Complexity, having the ability to choose the signs in a non-uniform way lets us choose a subset of pairs $(x,y)$ such that the minimizer $f$ over the subset is guaranteed to have (roughly) the same performance on the entire set. This allows us to obtain a guarantee for the distribution loss of a minimizer of a finite set of pairs chosen in an adaptive way to the family function $\F$. This set is in fact a core-set and the Komlos complexity helps us determine that possible size of a coreset. The tradeoff of considering a coreset vs.\ a random sample is that on one hand, it is clearly much simpler to obtain a random sample. On the other hand we will show in the following sections several examples for which a coreset can be significantly smaller than the random sample while maintaining the same guarantees. This will be done by showing that for a wide range of interesting problems in machine learning $K_m(\F) = o(R_m(\F))$.
\el{I like this explanation, I'm just not sure why it talks about pairs...}

\subsection{General Framework for Coresets}

In what follows it will be convenient to represent the function family as a family of vectors representing the functions. The vector family will be denoted as $Q$, where we use $q$ to denote a single vector corresponding to a function. We a universal function $f$ such that $f(x,q)$ corresponds to the value of the function matching $q$ when applied to $x$. We will assume from hereon that for any $x,q$ $|f(x,q)| \leq 1$; this allows to simplify the presentation. Our results can extend to $|f(x,q)| \leq R$ and will have dependencies on the bound $R$. Achieving these bounds is a purely technical task that we defer for the sake of readability.

Consider the function $F(q) = \sum_{i=1}^{m} f(x_i,q)$ and the signed sum of error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Now consider $F_1(q) = F(q) + E(q) = \sum_{i=1}^{m} f(x_i,q)  + \sum_{i=1}^{m} \sigma_i f(x_i,q)  = 2 \sum_{i | \sigma_i=1} f(x_i,q)$ and similarly $F_{-1}(q)$. We have that both $F_1(q), F_{-1}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|F_1(q) - F(q)| =  |E(q)|$.

%
The above is true for any choice of signs $\sigma_i$, specifically, for those minimizing $\max_q | E(q)|$.
Note that we can select signs such that $|E(q)| \le m K_m(\F)$.
A function is sketch-able if $K_m(\F) = o(m)$.
Typically we see that $K_m(\F) = O(c/m)$ or $K_m(\F)= c/\sqrt{m}$ for some $c$ which does not depend on $m$.

\begin{fact}
Let $\F$ be a function family with a corresponding Komlos Complexity $K_m(\F) = O(c/m)$. Let $F(q) = \sum_{i=1}^{n}f_i(q)$ where $f_i \in \F$. There exists a coreset for $F$ of size $k = O(c/\eps)$ whose error is at most $\eps n$.
\end{fact}
\begin{fact}

For any function family $\F$ with a corresponding Komlos Complexity $K_m(\F) = O(c/\sqrt{m})$ there exists a coreset of size 
$O(c^2/\eps^2)$ whose error is at most $\eps n$.

\end{fact}

\noindent Proving both facts is trivial and known in the folklore as ``the halving trick". 
At step one, create a coreset of size at most $n/2$ of items of weight $2$ and incur an error of at most $c$ or $c \sqrt{n}$.
Then, we repeat the process and create a coreset of size $n/4$ of items of weight $4$. Here, you incur error of $2c$ or $2c\sqrt{n/2}  = c\sqrt{2n}$.
Note that the sum of errors is a geometric sequence that is asymptotically dominated by its last element. 
Halting the compression at $\Theta(c/\eps)$ or $\Theta(c^2/\eps^2)$ items respectively achieves the goal.

\subsection{General Framework for Sketching}\label{sec:sketch}
We claim that low Komlos complexity implies concise streaming mergeable coresets. 
This section assumes the existence of an algorithm for finding low discrepancy signs $\sigma$.

\begin{theorem} \label{thm:streaming}
For any function family $\F$ with a corresponding Komlos Complexity with $K_m = O(c/m)$ there exists an fully-mergeable streaming coreset algorithms of size 
$$O\left(c\log^2\left(\min(\eps n/c, \log(n/\delta)) \right)/\eps\right)$$ 
whose error is at most $\eps n$ and failure probability at most $\delta$. For $\delta=0$ this results in a deterministic algorithm. The algorithm must be given $\eps, \delta$ as input and does not require a priori knowledge of the stream length $n$.
\end{theorem}

\begin{theorem} \label{thm:streaming2}
For any function family $\F$ with a corresponding Komlos Complexity of $K_m = O(c/\sqrt{m})$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
$$O\left(c^2\log^3\left(\min(\eps^2 n/c, \log(n/\delta)) \right)/\eps^2\right)$$ 
whose error is at most $\eps n$ and failure probability at most $\delta$. For $\delta=0$ this results in a deterministic algorithm. The algorithm must be given $\eps, \delta$ as input and does not require a priori knowledge of the stream length $n$.
\end{theorem}

The proofs of the above two theorems are quite technical and not very illuminating. Most of the ideas are already present in the recent breakthrough work on streaming quantiles \cite{DBLP:conf/focs/KarninLL16}.

@@

\section{A Simple Algorithm for Kernel Density Estimation}
From section \ref{sec:analytic} we know that the Komlos complexity of the Gaussian kernel is $K_m = O(\sqrt{d}/m)$. The result however is non-constructive in the sense that we are not aware of an algorithm that can obtain the $\sigma$ signs efficiently. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. We show that for any positive kernel $K_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section when $d > \sqrt{m}$. More importantly though, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = x_1,\ldots, x_n$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as 
$$ f(q) = \sum_{i=1}^{n} K(x_i,q) $$
Here, $K$ is a \emph{positive definite kernel} function, typically based on the distance between $x,y$. The most frequent examples include
$$ K(x,q) = \exp(- \|x-y\|_2^2/\lambda^2)\;\;\; K(x,q) = \exp(- \|x-q\|/\lambda) \; \mbox{and}\;\;\; K(x,y) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

@@

\begin{proof}
For any kernel $K$ there exist a mapping $\phi: \R^d \to {\cal V}$ to an inner product space $\cal V$ such that 
$$ K(x,q) = \ip{\phi(x), \phi(q)} $$
Using this function $\phi$ our objective function becomes
\[
\sum_{i=1}^m \sigma_i K(x_i,q) = \sum_{i=1}^m \sigma_i \ip{\phi(x_i), \phi(q)} =  \ip{ \sum_{i=1}^m \sigma_i \phi(x_i), \phi(q)} \leq  \|\phi(q)\| \cdot \left\|  \sum_{i=1}^m \sigma_i \phi(x_i) \right\| 
\]
Since $\|\phi(q)\| \leq 1$ we reduced the problem to bounding the norm of $ \sum_{i=1}^m \sigma_i \phi(x_i) $.
%
We will show by induction on $i$ that 
$$\left\| \sum_{j=1}^i \sigma_j \phi(x_j) \right\|^2 \le \sum_{j=1}^i \left\|\phi(x_j)\right\|^2 \leq i$$
This is trivially true for $i=1$ since $\|\phi(x)\| \leq 1$. 
Using our induction assumption we get
\begin{eqnarray*}
\left\| \sum_{j=1}^{i}\sigma_j \phi(x_j)\right\|^2 =& \left\|\sum_{j=1}^{i-1}\sigma_j \phi(x_j)\right\|^2 + \|\phi(x_i)\|^2 + 2\ip{ \sum_{j=1}^{i-1}\sigma_j \phi(x_j), \sigma_i \phi(x_i)} \leq & \text{induction assumption}\\
& \sum_{j=1}^{i-1} \|\phi(x_j)\|^2 + \|\phi(x_i)\|^2 + 2\sigma_i \sum_{j=1}^{i-1}\sigma_j K(x_j, x_i) = & \text{choice of } \sigma_i \\
& \sum_{j=1}^{i} \|\phi(x_j)\|^2 - 2|\sum_{j=1}^{i-1}\sigma_j K(x_j, x_i)| \le \sum_{j=1}^{i} \|\phi(x_j)\|^2 \\
\end{eqnarray*}
\end{proof}

@@

\bibliography{density}
